### 2012-12-14 doc/description.tex
new helpers updating mappings

$$
Application = (Class*, Instance*)
$$
$classes : Application \rightarrow Class*$ \\
$classes((classes)) = classes$ \\ \\
$instantiated : Class \times Application \rightarrow Boolean $
%TODO: variable types

@@

\subsection{Application Creation}
$$AddClass: Class \times Application \rightarrow Application $$

$
AddClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
{classes(a) \cup c}
$

@@

$
AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in properties(p) : name(q) \neq name(p)}{
 properties(c) \cup p }
$
$$AddAssociation : Label \times Association \times Application \rightarrow Application $$
$
AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(a) \cup s}
$

@@

\section{Database Component Model}
The relational database consists of database schema which defines the structure of the database and data. The database is a pair od sets:
$$
Database = (table*, data*)
$$

@@

\begin{center}
$ \rho : AppType \rightarrow DbType $ \\
$ \rho(APPSTRING) = DBSTRING $
\end{center}

@@

\begin{center}
$\rho : Class \rightarrow Table $ \\ 
$ \rho(c) = table(name(c), primaryKey("Id"), \rho(primitiveProperties(c), \varnothing $\\$ \cup \rho(associations(c)) \cup \rho(collectionProperties)$
\end{center}

@@

$
\rho : Property \times Database \rightarrow Column 
$

$\rho(p, d) = \begin{cases}
 \inference{cardinality(p) = 0 \\ \wedge \rho(owningClass(p)) \in classes(d)}{columns(\rho(owningClass(p))) \cup column(name(p), \rho(type(p)), \varnothing)} \\ \\

 \inference{cardinality(p) = 1 \\ \wedge \rho(owningClass(p) \in classes(d)}{columns(\rho(owningClass(p))) \cup column(name(p), \rho(type(p)), NOTNULL)} \\ \\

 \inference{cardinality(p) > 2 \\ \wedge \rho(owningClass(p) \in classes(d)}{\begin{gathered} table(name(p), primaryKey("Id"), column("value", \rho(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
 \end{cases}$

@@

$
\rho(a,d) = \begin{cases}
 \inference{startCardinality(a) = 0 \\ \wedge \rho(owningClass(p)) \in classes(d)  \wedge \rho(reference(a)) \in tables(d)}{
 \begin{gathered}
 foreignKeys(\rho(reference(a))) \\ \cup foreignKey(name(a), owningClass(a),  \varnothing) 
 \end{gathered}
 }

  \\ \\

 \inference{startCardinality(a) = 1 \\ \wedge \rho(owningClass(p) \in tables(d)  \wedge \rho(reference(a)) \in tables(d)}{\begin{gathered}
foreignKeys(\rho(reference(a))) \\ \cup  foreignKey(name(a), \\ owningClass(a),  NOTNULL)
\end{gathered}
} \\ \\

@@

%TODO FK nemusí mít jméno - odkazuje jen na tabulku a id

@@

\subsection{Creating of a System}
Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.

@@

$
\rho_{addClass}(c, d) = \inference{\forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
	\forall a \in \rho(associations(c)) : reference(a) \in tables(d) 
}{
\begin{gathered}
tables(d) \cup \rho(c) \cup \rho(collectionProperties(c)) \\ \cup \rho(associations(c))
\end{gathered}
}
$
\subsubsection{Add Property}
$
\rho_{addProperty}(l, p, d) = 
\inference{\exists t \in tables(d) : name(t) = l}{\rho(p,d)} 
$
\subsection{Add Association}
$
\rho_{AddAssociation}(l, a, d) = \inference{\exists t_1, t_2 \in tables(d) : name(t_1) = l \\ \wedge name(t_2) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_1) : name(f) = name(a) }{\rho(a,d)}
$

@@

\subsubsection{Remove Property}
$
\rho_{removeProperty}(l_c, l_p, d) = \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{
columns(t) \setminus c
} \\ \\
  \inference{}{}
 \end{cases}

$
%TODO 1:N M:N

>>>>

$$
Application = <Class*, Instance*>
$$
$classes : Application \rightarrow Class*$ \\
$classes(<cs, is>) = cs, cs \in Class* \wedge is \in Instance* $ \\ \\
$instantiated : Class \times Application \rightarrow Boolean $ \\
\\
$class : Label \times Application \rightarrow Class  $ \\ 
$class(l, a) = c, c \in classes(a) \wedge name(c) = l, l \in Label, a \in Application* $ 
%TODO: variable types

@@

\subsection{Application Creation}
$$addClass: Class \times Application \rightarrow Application $$

$
addClass(c, a) = \inference{\forall s \in classes(A) : name(s) \neq name(c)}
{classes(a) \cup c}
$

@@

$
AddProperty(l, p, a) = \inference{\exists c \in classes(a) : name(c) = l \\ \wedge \forall q \in properties(p) : name(q) \neq name(p)}{
 properties(c) \cup p }
$
$$AddAssociation : Label \times Association \times Application \rightarrow Application $$
$
AddAssociation(l, s, a) = \inference{\forall f \in associations(a) : name(f) \neq name(s) \\ \wedge \exists c \in classes(a) : name(c) = l \\ \wedge \exists k \in classes(a) : name(k) = reference(s) }{associations(c) \cup s}
$

@@

\section{Database Component Model}
The relational database consists of database schema which defines the structure of the database and data. The database is a pair od sets:
$$
Database = (table*, data*)
$$
$table : Label \times Database \rightarrow Table $ \\
$table(l, d) = t, t \in tables(d) \wedge name(t) = l$ \\

@@

\begin{center}
$ \rho_{t} : AppType \rightarrow DbType $ \\
$ \rho_{t}(APPSTRING) = DBSTRING $
\end{center}

@@

\begin{center}
$\rho_{c}: Class \rightarrow Table $ \\ 
$\rho_{c}(c) = table(name(c), primaryKey("Id"), \rho_{p}(primitiveProperties(c), \varnothing $\\$ \cup \rho_{a}(associations(c)) \cup \rho_{p}(collectionProperties)$
\end{center}

@@

$
\rho_p : Property \times Database \rightarrow Column 
$

$\rho_p(p, d) = \begin{cases}
 \inference{cardinality(p) = 0 \\ \wedge \rho_c(owningClass(p)) \in tables(d)}{\begin{gathered}
 columns(table(name(owningClass(p)),d) \\ \cup column(name(p), \rho_t(type(p)), \varnothing)
\end{gathered}
} \\ \\

 \inference{cardinality(p) = 1 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{columns(t) \cup column(name(p), \rho_t(type(p)), NOTNULL)} \\ \\

 \inference{cardinality(p) > 2 \\ \wedge \exists t : t = table(name(owningClass(p)), d)}{\begin{gathered} table(name(p), primaryKey("Id"), column("value", \rho_t(type(p))), \\ foreignKey(name(p), name(owningClass(p))) \end{gathered}}
 \end{cases}$

@@

$
\rho_a(a,d) = \begin{cases}
 \inference{startCardinality(a) = 0 \\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{
 \begin{gathered}
 foreignKeys(u) \\ \cup foreignKey(name(a), name(t),  \varnothing) 
 \end{gathered}
 }
  \\ \\
 \inference{startCardinality(a) = 1 \\\\ \wedge \exists t : t = table(name(owningClass(p)), d) \\ \wedge \exists u : u = table(reference(a)), d)}{\begin{gathered}
foreignKeys(u) \\ \cup  foreignKey(name(a), name(t),  NOTNULL)
\end{gathered}
} \\ \\

@@

%TODO handle instances

@@

%TODO jak se postupně vyvíjí systém
\subsection{Creating of a System}
Creational operations does not have impact on stored data they change the structure of the application and database by adding new objects into model.

@@

$
\rho_{addClass}(c, d) = \inference{\forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
}{
\begin{gathered}
tables(d) \cup \rho(c) \cup \rho(collectionProperties(c)) \\ \cup \rho(associations(c))
\end{gathered}
}
$
\subsubsection{Add Property}
$
\rho_{addProperty}(p, d) = 
\inference{\exists t \in tables(d) : name(t) = name(owningClass(p))}{\rho(p,d)} 
$
\subsection{Add Association}
$
\rho_{AddAssociation}(a, d) = \inference{\exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) }{\rho(a,d)}
$

@@

\subsubsection{Remove Property}
$
\rho_{removeProperty}(l_c, l_p, d) = \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{
columns(t) \setminus c
} \\ \\
  \inference{\exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{tables(d) \setminus u }
 \end{cases}
$

\subsubsection{Remove Association}
$
\rho_{removeAssociation}(l_c, l_a, d) = \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = l_c \wedge }{foreignKeys(\rho(reference(a))) \ }

 \inference{}{}
 \end{cases}
$
